---
title: "Pakistan Drone Attacks Analysis"
output:
  html_document: default
  html_notebook: default
---

## 1. Loading The Dataset And Required Imports
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(scales)
setwd("E:/MOOC")
attacks <- read.csv("PakistanDroneAttacks.csv",stringsAsFactors=FALSE)
```

## 2. Dimensions Of Dataset
```{r}
dim(attacks)
```

## 3. Description of the variables and thier types.
```{r}
str(attacks)
```

## 4. Data Cleaning 
As part of data cleanning, I have performed the foolowing operations on the dataset.
1. Converted the given date format to standard date format.
2. Added year column to dataset which can be used for timeline analysis.
3. Converted the city and province variables to factor variables.
4. Added tenure variable for tenure analysis.
```{r}
attacks$Date <- as.Date(attacks$Date,'%A, %B %d, %Y')
attacks$Year <- format(attacks$Date,'%Y')
attacks$City <- as.factor(attacks$City)
attacks$Province <- factor(attacks$Province)
attacks$tenure <- ifelse(attacks$Year <= 2008,"George W. Bush","Barack Obama") 
str(attacks)
```


```{r}
ggplot(subset(attacks,!is.na(Year)),aes(Year,group=1)) + 
geom_freqpoly(stat = "count",size=0.5,color="red")+
geom_point(stat = "count")+  
scale_y_continuous(breaks = seq(0,120,10))+  
labs(x="Year",y="Number of Attacks",title="                                         Timeline Of Number Of Attacks")+
theme_bw()
```


