---
title: "Pakistan Drone Attacks Analysis"
output:
  html_document: default
  html_notebook: default
---

## 1. Loading The Dataset And Required Imports
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(scales)
setwd("E:/MOOC")
attacks <- read.csv("PakistanDroneAttacks.csv",stringsAsFactors=FALSE)
```

## 2. Dimensions Of Dataset
```{r}
dim(attacks)
```

## 3. Description of the variables and thier types.
```{r}
str(attacks)
```

## 4. Data Cleaning 
* As part of data cleaning, I have performed the following operations on the dataset.
 + Converted the given date format to standard date format.
 + Added year column to dataset which can be used for timeline analysis.
 + Converted the city and province variables to factor variables.
 + Added tenure variable for tenure analysis.
 + Removed the "Special.Mention..Site.", "Comments", "References", "Longitude", "Latitude", "Temperature.C.",  and "Temperature.F." columns, as they    won't be much useful for my exploration. 
```{r}
attacks$Date <- as.Date(attacks$Date,'%A, %B %d, %Y')
attacks$Year <- format(attacks$Date,'%Y')
attacks$Province <- factor(attacks$Province)
attacks$Tenure <- ifelse(attacks$Year <= 2008,"George W. Bush","Barack Obama")
attacks <- subset(attacks, select = -c(Comments,References,Temperature.C., Temperature.F., Special.Mention..Site.))

# Removing NA Values
attacks$Al.Qaeda[is.na(attacks$Al.Qaeda)] <- 0
attacks$Taliban[is.na(attacks$Taliban)] <- 0
attacks$Civilians.Min[is.na(attacks$Civilians.Min )] <- 0
attacks$Civilians.Max[is.na(attacks$Civilians.Max)] <- 0
attacks$Foreigners.Min[is.na(attacks$Foreigners.Min)] <- 0
attacks$Foreigners.Max[is.na(attacks$Foreigners.Max)] <- 0
attacks$Total.Died.Min[is.na(attacks$Total.Died.Min)] <- 0
attacks$Total.Died.Mix[is.na(attacks$Total.Died.Mix)] <- 0
attacks$Injured.Min[is.na(attacks$Injured.Min)] <- 0
attacks$Injured.Max[is.na(attacks$Injured.Max)] <- 0

str(attacks)
```

## 5. Overall Metrics
In this part of exploration lets explore the total number casualties occurred till date:
```{r}
total_strikes <- sum(attacks$No.of.Strike)
total_alqaeda <- sum(attacks$Al.Qaeda)
total_taliban <- sum(attacks$Taliban)
total_civilians <- sum(attacks$Civilians.Max)
total_foreigners <- sum(attacks$Foreigners.Max)
total_injured <- sum(attacks$Injured.Max)
total_deaths <- sum(attacks$Total.Died.Mix)
total_Women_Children <- sum(attacks$Women.Children == 'Y')

sprintf("Total Number Of Strikes : %s",total_strikes)
sprintf("Total Number Of AL.Qaeda Terrorists Killed: %s",total_alqaeda)
sprintf("Total Number Of Taliban Terrorists Killed: %s",total_taliban)
sprintf("Total Number Of Civilian Casualties: %s",total_civilians)
sprintf("Total Number Of Foreigners killed: %s",total_foreigners)
sprintf("Total Number Of Injured: %s",total_injured)
sprintf("Total Number Of Deaths: %s",total_deaths)
sprintf("Total Number Of Strikes That Included Women and Children: %s",total_Women_Children)
```
The overall metrics show that a lot of civilian casualties were reported. The civilian casualties contribute to more than 67% of the total deaths. 

## 6.Timeline Analysis
Now lets explore the number of drone strikes carried out over year timeline. 
```{r}
ggplot(attacks,aes(Year,group=1)) + 
geom_line(stat = "summary",fun.y=sum,size=0.5,color="red",aes(y=No.of.Strike))+
geom_point(stat = "summary",fun.y=sum,aes(y=No.of.Strike))+  
scale_y_continuous(breaks = seq(0,130,10))+  
labs(x="Year",y="Number of Attacks",title="Timeline Of Number Of Drone strikes Carried Out")+
theme_bw()
```

The above timeline shows that maximum number of attacks were carried out b/w 2008 and 2015. Now lets explore the civilian casualties and number of terrorists killed over a year timeline.
```{r fig.width = 10, fig.height = 5}
cols <- c("Civilians"="blue","Terrorits"="red")
ggplot(attacks,aes(x=Year,group=1))+
  geom_line(stat = "summary",fun.y=sum,size=0.5,aes(y=Civilians.Max,color="Civilians"))+
  geom_point(stat = "summary",fun.y=sum,aes(y=Civilians.Max))+
  geom_line(stat = "summary",fun.y=sum,size=0.5,aes(y=Al.Qaeda+Taliban,color="Terrorits"))+
  geom_point(stat = "summary",fun.y=sum,aes(y=Al.Qaeda+Taliban))+
  labs(x="Year",y="Number Of Casualties Reported",title="Timeline Comparison Of Terrorists and Civilian Casualties")+
  scale_colour_manual(name="Civilians V/S Terrorists Killed",values=cols)+
  scale_y_continuous(breaks = seq(0,1000,100))+
  theme_bw()
```

The Timeline analysis shows that, till 2012 the number of civilian casualties reported were high when compared to the number of terrorists killed. Now lets explore the number of people injured over a year timeline.
```{r}
ggplot(attacks,aes(x=Year,group=1))+
  geom_line(stat="summary",fun.y = sum,aes(y=Injured.Max),color= "blue")+
  geom_point(stat="summary",fun.y = sum,aes(y=Injured.Max))+
  scale_y_continuous(breaks = seq(0,500,50))+
  labs(x="Year",y="Number of Injured People",title="Timeline Comparison of Injured People")+
  theme_bw()
```

 + The above timeline shows that most number of people were injured in b/w 2008 and 2015, which was the period in which most of the drone strikes were carried out. Now lets take a look at province variable in the dataset, to see which province of Pakistani was targeted the most.

## 7. Province And City Based Exploration.
```{r}
ggplot(attacks,aes(x=Province))+
  geom_bar(width = 0.3,fill="#A52A2A")+
  geom_text(stat='count',aes(label=..count..),vjust=-1)+
  scale_y_continuous(breaks=seq(0,500,50),limits = c(0,450))+
  labs(x="Year",y="Number Of Attacks Carried Out",title="Province Wise Attacks")+
  theme_bw()
```

The Province wise analysis show that most of the attacks were carried out in FATA province because its is a tribal region in Pakistan and its not under much control of Pakistan government and many terrorist groups carry out their operations from this region. Now lets explore the city variable to see which cities were targeted the most.
```{r}
table(attacks$City)
```

The City variable seems to have duplicates of the values "South Waziristan", Lets rename the duplicate occurrences, so that they point to a single value.  
```{r}
attacks$City <- ifelse(tolower(attacks$City) == "south waziristan","South Waziristan",attacks$City)
table(attacks$City)
```


```{r}
mylog10_trans <- function (base = 10) 
{
  trans <- function(x) log(x + 1, base)
  inv <- function(x) base^x
  trans_new(paste0("log-", format(base)), trans, inv, log_breaks(base = base), 
            domain = c(1e-100, Inf))
}

ggplot(subset(attacks,attacks$City != ""),aes(City))+
  geom_bar(fill="#A52A2A")+
  geom_text(stat='count',aes(label=..count..),hjust=-0.7)+
  scale_y_continuous(trans = "mylog10",limits = c(0,400))+
  labs(x="City Attacked",y="No Of Attacks Carried Out",title="City Wise Attacks")+
  coord_flip()+
  theme_bw()
```

 * The City analysis shows that, most attacked cities are north and south waziristan. These two cities belong to the FATA province that's the reason why they are targeted the most.
 


