---
title: "Pakistan Drone Attacks Analysis"
output:
  html_document: default
  html_notebook: default
---

## 1. Loading The Dataset And Required Imports
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(scales)
setwd("E:/MOOC")
attacks <- read.csv("PakistanDroneAttacks.csv",stringsAsFactors=FALSE)
```

## 2. Dimensions Of Dataset
```{r}
dim(attacks)
```

## 3. Description of the variables and thier types.
```{r}
str(attacks)
```

## 4. Data Cleaning 
* As part of data cleaning, I have performed the following operations on the dataset.
 + Converted the given date format to standard date format.
 + Added year column to dataset which can be used for timeline analysis.
 + Converted the city and province variables to factor variables.
 + Added tenure variable for tenure analysis.
 + Removed the "Special.Mention..Site.", "Comments", "References", "Longitude", "Latitude", "Temperature.C.",  and "Temperature.F." columns, as they    won't be much useful for my exploration. 
```{r}
attacks$Date <- as.Date(attacks$Date,'%A, %B %d, %Y')
attacks$Year <- format(attacks$Date,'%Y')
attacks$City <- as.factor(attacks$City)
attacks$Province <- factor(attacks$Province)
attacks$Tenure <- ifelse(attacks$Year <= 2008,"George W. Bush","Barack Obama")
attacks <- subset(attacks, select = -c(Comments,References,Longitude,Latitude,Temperature.C., Temperature.F., Special.Mention..Site.))

# Removing NA Values
attacks$Al.Qaeda[is.na(attacks$Al.Qaeda)] <- 0
attacks$Taliban[is.na(attacks$Taliban)] <- 0
attacks$Civilians.Min[is.na(attacks$Civilians.Min )] <- 0
attacks$Civilians.Max[is.na(attacks$Civilians.Max)] <- 0
attacks$Foreigners.Min[is.na(attacks$Foreigners.Min)] <- 0
attacks$Foreigners.Max[is.na(attacks$Foreigners.Max)] <- 0
attacks$Total.Died.Min[is.na(attacks$Total.Died.Min)] <- 0
attacks$Total.Died.Mix[is.na(attacks$Total.Died.Mix)] <- 0
attacks$Injured.Min[is.na(attacks$Injured.Min)] <- 0
attacks$Injured.Max[is.na(attacks$Injured.Max)] <- 0

str(attacks)
```

## 5. Overall Metrics
In this part of exploration lets explore the total number casualties occurred till date:
```{r}
total_strikes <- sum(attacks$No.of.Strike)
total_alqaeda <- sum(attacks$Al.Qaeda)
total_taliban <- sum(attacks$Taliban)
total_civilians <- sum(attacks$Civilians.Max)
total_foreigners <- sum(attacks$Foreigners.Max)
total_deaths <- sum(attacks$Total.Died.Mix)
total_Women_Children <- sum(attacks$Women.Children == 'Y')

sprintf("Total Number Of Strikes : %s",total_strikes)
sprintf("Total Number Of AL.Qaeda Terrorists Killed: %s",total_alqaeda)
sprintf("Total Number Of Taliban Terrorists Killed: %s",total_taliban)
sprintf("Total Number Of Civilian Casualties: %s",total_civilians)
sprintf("Total Number Of Foreigners killed: %s",total_foreigners)
sprintf("Total Number Of Deaths: %s",total_deaths)
sprintf("Total Number Of Strikes That Included Women and Children: %s",total_Women_Children)
```
The overall metrics show that a lot of civilian casualties were reported. The civilian casualties contribute to more than 67% of the total deaths. Now lets compare civilian casualties to number of terrorists killed over a year timeline.  

## 6.Timeline Analysis
```{r fig.width = 10, fig.height = 5}
cols <- c("Civilians"="blue","Terrorits"="red")
ggplot(attacks,aes(x=Year,group=1))+
  geom_line(stat = "summary",fun.y=sum,size=0.5,aes(y=Civilians.Max,color="Civilians"))+
  geom_point(stat = "summary",fun.y=sum,aes(y=Civilians.Max))+
  geom_line(stat = "summary",fun.y=sum,size=0.5,aes(y=Al.Qaeda+Taliban,color="Terrorits"))+
  geom_point(stat = "summary",fun.y=sum,aes(y=Al.Qaeda+Taliban))+
  labs(x="Year",y="Number Of Casualties Reported",title="Timeline Comparison Of Terrorists and Civilian Casualties")+
  scale_colour_manual(name="Civilians V/S Terrorists Killed",values=cols)+
  scale_y_continuous(breaks = seq(0,1000,100))+
  theme_bw()
```

The Timeline analysis shows that, till 2012 the number of civilian casualties reported were high when compared to the number of terrorists killed. Now lets explore the number of drone strikes carried out over year timeline. 
```{r }
ggplot(attacks,aes(Year,group=1)) + 
geom_freqpoly(stat = "summary",fun.y=sum,size=0.5,color="red",aes(y=No.of.Strike))+
geom_point(stat = "summary",fun.y=sum,aes(y=No.of.Strike))+  
scale_y_continuous(breaks = seq(0,130,10))+  
labs(x="Year",y="Number of Attacks",title="Timeline Of Number Of Drone strikes Carried Out")+
theme_bw()
```





